---
title: "Bayesian IRT"
author: "Reginald Ziedzor"
date: "4/29/2020"
output: rmarkdown::github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bayesian Item Response Theory using Stan

This is an introduction to using Stan for item score analysis. We are using the LSAT dataset from the `ltm` package in R. 

```{r}
data(LSAT, package = "ltm")
```

We can look at the first few observations of the LSAT dataset
```{r}
head(LSAT)
```

## Running Stan Code


```{r}
library(rstan)
options(mc.cores = parallel::detectCores()-1)
rstan_options(auto_write = TRUE)
```
Using `rstan` to run the Stan code.

```{stan, output.var = "IRT_1PL", cache = TRUE}
data {
  int N_I; //number of items
  int N; //number of examinees
  int<lower = 0, upper = 1> Y[N, N_I]; //observed data
}

parameters {
  vector[N] theta; //ability levels of examinees
  vector[N_I] diff; //item difficulty
  real<lower = 0> disc; //ensuring item discrimination parameter is nonnegative
  real<lower = 0> sigma_disc;
  real mu_diff;
}

model {
  real p_ij; //probability of correctly responding to an item
  
  for (j in 1:N) {
    for(i in 1:N_I) {
      
      p_ij = disc*(theta[j] - diff[i]);
      
      Y[j,i] ~ bernoulli_logit(p_ij); 
    }
  }
  //Priors
  mu_diff ~ normal(0,2);
  theta ~ std_normal();
  diff ~ normal(mu_diff,5);
  disc ~ normal(0,sigma_disc);
  sigma_disc ~ cauchy(0,2); 
}

```

Posterior Sampling
```{r, results='hide', message=FALSE, cache=TRUE}

mod <- rstan::sampling(IRT_1PL, chains = 3, data = list(N = nrow(LSAT), N_I = ncol(LSAT), Y = LSAT))
```

Posterior Estimates
```{r}
print(mod, pars = c("disc", "diff"))
```

## Diagnostic Plots 
```{r, fig-margin, fig.margin=TRUE}
rstan::traceplot(mod, pars = c("disc", "diff"))
rstan::stan_dens(mod, pars = c("disc", "diff"))

```

## Frequentist Parameter Estimation 
Using the `ltm` package to obtain maximum likelihood estimates
```{r}
library(ltm)
model_rasch <- rasch(data = LSAT, IRT.param = TRUE)
summary(model_rasch)
```


